---
title: "NSW-500"
author: "Paris Heard" 
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Learning Objectives

-   Understand the relationship between reported effective collaboration and remote work
-   Explore the relationship between leadership, team culture, and the effectiveness of collaboration
-   Gain insight into the desires of the average employee, by industry

## Package Installations

```{r}
set.seed(2)

library(tidyverse)
library(ggplot2)
library(dplyr)
library(infer)
library(HH)
library(likert)
library(foreign)
library(colorspace)
library(RColorBrewer)
library(grid)
library(knitr)
```

## Creating Data-frame

```{r cars}
sdata <- read.csv("2021_rws.csv", header = TRUE, sep = ",")
```

## Cleaning Data

```{r}
head(sdata)
```

Crafting a new data frame for only those employees who have worked remotely in the past six months, and removing all ID numbers.

```{r}
recent_RW <- sdata %>%
  filter(worked_RW_past_six_months == "Yes") %>%
  dplyr::select(birth_year, gender, yr_experience, household_size, metro_regional, frequency_RW, perc_required_onsite, six_months_collaborate_easily_RW, hybrid_biggest_barriers_one, hybrid_smallest_barriers_one, feel_better_when_RW, feel_better_seeing_colleagues_onsite, more_active_when_RW, team_works_well_RW, productivity_RW, easy_to_contact_employees_RW) %>%
  mutate(across(where(~all(str_detect(.,"%"))), parse_number))

recent_RW
```

## Fitting Regression Models

### Remote Work Frequency & Team Collaboration

Does the frequency of remote work relate to perceived ability to collaborate among a team? Let's fit a regression model to observe the relationship.

```{r}
collab_frq_model <- lm(six_months_collaborate_easily_RW ~ frequency_RW, data = recent_RW)
get_regression_table(collab_frq_model)
```

In order to visualize the collaboration responses in likert format, we will need to do a bit of data wrangling. Let's select our target variables and pivot likert responses to a wider format.

```{r}
collab_freq <- recent_RW %>%
  dplyr::select(six_months_collaborate_easily_RW, frequency_RW) %>%
  group_by(six_months_collaborate_easily_RW, frequency_RW) %>%
  pivot_wider(names_from = "six_months_collaborate_easily_RW", 
              values_from = "six_months_collaborate_easily_RW", 
              values_fn = function(x) sum(!is.na(x)), values_fill = 0) %>%
  arrange(desc(frequency_RW))
```

```{r}
# likert1 <- likert(frequency_RW~., collab_freq, 
#                 ReferenceZero=3, 
#                 ylab = "Frequency of Remote Work", 
#                 main = list("Collaboration Success", x = unit(.62, "npc")), 
#                 auto.key = list(columns = 2, reverse.rows = T))
likert1 <- HH::likert(frequency_RW~., collab_freq, ReferenceZero=3, ylab = "Remote Work Frequency", main =
list("Ease of Remote Collaboration", x=unit(.62, "npc")), auto.key = list(columns = 2,
reverse.rows = T))

likert1

png("../NSW-500/output/likert_1.png", height=720, width=1080)

likert1
dev.off()


```
