---
title: "NSW-500"
author: "Paris Heard" 
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Learning Objectives

-   Understand the relationship between reported effective collaboration and remote work
-   Explore the relationship between leadership, team culture, and the effectiveness of collaboration
-   Gain insight into the desires of the average employee, by industry

## Package Installations

```{r}
set.seed(2)

library(tidyverse)
library(ggplot2)
library(dplyr)
library(infer)
```

## Creating Data-frame

```{r cars}
sdata <- read.csv("2021_rws.csv", header = TRUE, sep = ",")
```

## Cleaning Data

```{r}
head(sdata)
```

Crafting a new data frame for only those employees who have worked remotely in the past six months, and removing all ID numbers.

```{r}
recent_RW <- sdata %>%
  filter(worked_RW_past_six_months == "Yes") %>%
  select(birth_year, gender, yr_experience, household_size, metro_regional, frequency_RW, perc_required_onsite, six_months_collaborate_easily_RW, hybrid_biggest_barriers_one, hybrid_smallest_barriers_one, feel_better_when_RW, feel_better_seeing_colleagues_onsite, more_active_when_RW, team_works_well_RW, productivity_RW, easy_to_contact_employees_RW)

recent_RW
```

```{r}
recent_RW_tidy <- recent_RW %>%
  pivot_longer()
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
